Rebuttal JMLR 14-133

### 2014-09-05 pkg/partykit/inst/mloss/review2/p2p.tex
elaborate based on discussion with AZ
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@597 255f4249-2263-4417-b1ea-8498779e27a7

\textit{
This is a clean and concise description of an interesting set of tools for
data analysis with recursive partitioning.  It is mainly a description or
announcement, and has minimal polemic or pedagogical value.  It will be a
useful reference for those a basic outline of the software's capabilities,
and provides a useful bibliography.}

Tja.

mention: previous round of reviews asked for this format of the paper

@@

yes

maybe worth mentioning: due to the flexible infrastructure, the two
specialized methods (ctree, mob) have become more flexible compared to
earlier implementations (especially the mob framework)

@@

\textit{Quibbles and queries are:}
\begin{enumerate}
\item 
\textit{
The legend for Table 1 begins: ``Selected tree algorithms than can be
interfaced through partykit...''  Surely it is the output from selected
implementations that can be interfaced, not the algorithm used.  The word
`algorithm' seems to me anyway inappropriate in this context.  The same (or
very similar) algorithm may have different implementations.  Different
methods with their inevitably different algorithms (including, in partykit:
classical `constant' trees, model-based trees, and conditional inference
trees) further wide widen the range of implementations.}

rephrase: interfaces data structures used by different implementations of different
tree growing methods / algorithms.

\item 
\textit{
The distinction between implementation (what the authors call `algorithm')
and output is crucial also because as far as I am able to judge, partykit's
abilities are limited to taking over and processing some part only of the
output from, e.g., R's rpart package.  As I understand, there is no ability
to use or process, e.g., results from rpart's on the fly cross-validation. 
Would it be straightforward to incorporate structure that is able to
accommodate such information?  See also point 3 that now follows.}

Hm, maybe store such information in \texttt{info} slots?

@@

See \citep{Hothorn+Hornik+Zeileis:2006}.

But probably also rephrase...

@@

Work in progress.

@@

infrastructure independent of such things.

@@

completely.

@@

pfff

@@

mention: previous round of reviews asked for this format

@@

not a big software project, just two main developers, hence sometimes longer breaks

we feel that this is not an indicator of the quality of the code, though

@@

E-mails to the maintainer address were always possible and used fairly regularly
by both end-users and other package developers. A mailing list was added now.

@@

Hard to judge, but reverse depends: NHEMOtree, quint. Number of citations of party paper.

\textit{
Support for trees produced by other libraries is provided, but I am not
certain how appealing the infrastructure will be to developers who want to
represent trees from other systems.  It seems like for non-PMML formatted
trees, implementing a conversion layer to partykit objects wouldn't be any
easier or harder than other packages.}

depends, eh?

examples: CTree and display of p-values, J4.8 and multiway splits

\textit{
Clearly, for developing new tree algorithms, though, partykit would be a good choice.}

We actually have proof!

@@

schwierig.

@@

schwierig.

@@

done.

>>>>

\textit{
This is a clean and concise description of an interesting set of tools for
data analysis with recursive partitioning.  It is mainly a description or
announcement, and has minimal polemic or pedagogical value.}

When preparing this revision of our earlier version JMLR-07-240, we closely
followed the suggestions given in the corresponding reviews. Our
understanding of this remark of Referee 3 on JMLR-07-240
\begin{quote} 
Surely it is possible to give broadbrush information. There should be direct
reference to web-based
materials that give more detailed information and examples.
\end{quote}
was that the four page description should really focus on 
broadbrush informtion, with all the interesting details being explained and
discussed in the packages's documentation.

\textit{It will be a
useful reference for those a basic outline of the software's capabilities,
and provides a useful bibliography.}

Thank you.

@@

Thank you for this nice and concise summary of the four page description.

When we started working on partykit, we redesigned all classes from the
earlier party package from scretch and also reimplemented ctree and mob
without much overlap between the party and partykit code bases. As a
consequence of the much cleaner partykit design, the code base for both
ctree and mob could be reduced considerably while at the same time making
both implementations more powerful due to more flexible options.

@@

Thank you.

\textit{Quibbles and queries are:}
\begin{enumerate}
\item 
\textit{
The legend for Table 1 begins: ``Selected tree algorithms than can be
interfaced through partykit...''  Surely it is the output from selected
implementations that can be interfaced, not the algorithm used.}

Yes, thank you. We completely agree with you on the distinction between
`algorithm' and it's `implementation' and changed the caption accordingly.

\textit{The word
`algorithm' seems to me anyway inappropriate in this context.  The same (or
very similar) algorithm may have different implementations.  Different
methods with their inevitably different algorithms (including, in partykit:
classical `constant' trees, model-based trees, and conditional inference
trees) further widen the range of implementations.}

We completely agree and consistently use "algorithm" for the concept
and "implementation" for the software throughout the paper now.

\item 
\textit{
The distinction between implementation (what the authors call `algorithm')
and output is crucial also because as far as I am able to judge, partykit's
abilities are limited to taking over and processing some part only of the
output from, e.g., R's rpart package.  As I understand, there is no ability
to use or process, e.g., results from rpart's on the fly cross-validation.}

Yes, this is right. It is not possible to directly work on the rpart
results.  The reason is that we coerce rpart objects to constparty objects
and this class does not capture cost-complexity pruning since not all
algorithms have a notion of a risk function and thus perform cost-complexity
pruning.  The tools provided by partykit require that pruning is performed
by rpart or J4.8 or by user-implemented add-ons.

\textit{Would it be straightforward to incorporate structure that is able to
accommodate such information?  See also point 3 that now follows.}

Yes, one `just' needs to define a new class that inherits from constparty,
possibly containing a cost-complexity table in the info slot with
appropriate methods for pruning (the nodeprune function can be used).  An
example on the use of nodeprune is given in the constparty vignette,
Section~3, and mob has a novel internal pruning functionality.

@@

This is a somewhat philosphical issue, discussed in more detail in
Section~3.4 of \cite{Hothorn+Hornik+Zeileis:2006}. If one grows a tree
on a very large data set with nominal level of 5\%, many statistically
significant but practically irrelevant splits will be found. For small and
medium sized problems, conditional inference trees with nominal level of 5\% 
and appropriate adjustment for multiplicity were shown to produce trees
of the correct size \citep[simulation results
in][]{Hothorn+Hornik+Zeileis:2006}.

@@

In principle it is straightforward to set-up a class for forests as a list
of partynode objects. The advantage is that only the tree
structures vary across the ensemble but not the data. However, we found such
a class of rather limited use because we didn't see a unified way to produce
meaningful plots (variable importances, partial dependencies etc) for
ensembles obtained by different forest algorithms yet. Implementing a new
class for forests (maybe called parties or orgie) is on our to do list.

@@

The general infrastructure provided by partykit covers the representation of
trees but not tree induction itself. This makes sense because the details of
how splits are searched for and selected differ considerably between
algorithms. The same is true with respect to finding the right-sized tree.

The implementing of tree algorithms is very much simplified by partykit
because the programmer can concentrate on the technical details of learning
the tree while using the partykit infrastructure for representing and
further processing the resulting tree. The constparty vignette contains a
very dense example of such an implementation (Section 3). Note that a
non-trivial tree algorithm can be implemented by three small R functions
only.

We also added an example showing how the partykit infrastructure can be used
to quickly implement model assessment strategies.

@@

Completely. <Z FIXME> mob.

@@

see additions to constparty vignette
<Z FIXME>

@@

Our understanding of the reviews on an earlier version of this paper was
that the referees asked for this format, please see our answer to Referee 1.

@@

The package was written by only two developers, hence sometimes longer
breaks to the new jobs or family businesses could not be avoided.
It is true that there was not much development in partykit between 2009 and
2013. During this time, two offspring packages (evtree and psychotree) where
developed instead and partykit was actively maintained. Please note that the
development of these two new packages did only bring some minor design or
implementation deficits in partykit to light. This raised our confidence in
partykit. For these reasons we feel that pure activity this is not an indicator of 
the quality of the code.

@@

E-mails to the maintainer address were always possible and are in fact the standard
way of communicating with CRAN package maintainers (and developers) and were 
used fairly regularly by both end-users and other package developers. 
Nevertheless, a mailing list was added now.

@@

One can argue about this statement being close to reality. As of now, the
package website lists seven reverse dependencies on CRAN, we are not
involved in three of them. Only 6\% of all CRAN packages have more 
reverse dependencies than partykit. And we hope that some of the 24 
reverse dependencies of party will switch to partykit
(only 2\% have more reverse dependencies than party) now that the reference
implementation of conditional inference trees is the one from partykit.
The ctree-publication was cited 643 times and the mob-paper 103 times 
(Google scholar), so we are confident that people will increasingly use
either ctree or mob and also hopefully learn about the partykit
infrastructure this way.

\textit{
Support for trees produced by other libraries is provided, but I am not
certain how appealing the infrastructure will be to developers who want to
represent trees from other systems.  It seems like for non-PMML formatted
trees, implementing a conversion layer to partykit objects wouldn't be any
easier or harder than other packages. Clearly, for developing new tree
algorithms, though, partykit would be a good choice.}

Our experience while implementing the original code for ctree 10 years ago
was that the infrastructure existing then had way too many
algorithm-specific things hardwired than to be of any use. While rpart
always allowed to change risk functions, doing something outside of the
CART-way of finding splits was not possible. Also printing and plotting
facilities were not able to, for example, put p-values onto inner nodes.

Our party implementation of ctree suffered the same problems, we had to work
extremely hard to somehow use the party plot methods to plot J4.8 multiway 
trees fitted by the RWeka package.

These experiences led to the partykit package. The reimplementation of
ctree, mob and the new implementations of trees in evtree and psychotree
turned out to be rather straightforward and we personally did benefit a lot
from the partykit infrastructure already. We suppose this is also true for
the developers of the NHEMOtree and quint packages.

@@

Z

@@

Z

Total cost of ownership

demo mit mlbench:Satellite und rpart und J4.8

@@

Based on our experience with the party package for 10 years now, the fact
that other developers started packages using partykit, and the feedback via
email and on conferences, most important the ``Workshop on Classification
and Regression Trees'' that took place in March 2014, we are more optimistic
about the popularity of partykit.
